NLIZE: a Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-27 paper/background.tex
updates on background
update attention illustration
update task

\subsection{Natural Language Inference}
\label{sec:languageInference}
Natural Language Inference (NLI)~\cite{DaganRothSammons2013} is an important machine understanding task in NLP.
The goal of NLI is to predict the relationship between a premise (\textbf{P}) sentence and a hypothesis (\textbf{H}) sentence. 
The prediction falls in one of three categories: \emph{entailment} (\textbf{E}), \emph{contradiction} (\textbf{C}), \emph{neutral} (\textbf{N}).
A simple example is shown in Table~\ref{tab:NLI}.
In this case, the premise is ``Jim ate an apple''. 
The hypothesis statement ``Jim ate fruit.'' can be concluded from the premise. Therefore, the relationship between premise and hypothesis is \emph{entailment}. However, we should note that such a relationship is not necessarily reversible. Since the concept ``fruit'' is less restrictive than that of ``apple'', we can not conclude ``Boy ate an apple'' from the statement ``Kid ate fruit''. The same logic applies to the hypothesis of ``Jim ate a Fuji apple.'', in which the premise neither imply nor oppose the hypothesis, therefore, their relationship is \emph{neutral}.
%\emph{Contradiction} means the textual of hypothesis opposes that of premise;
%and \emph{Neutral} implies no conclusion can be made on either \emph{Entailment} or \emph{Contradiction}.

\begin{table}[htbp]
\label{tab:NLI}
\centering
\caption{An illustration of natural language inference.}
 \begin{tabular}{c | c c c c} 
 \hline
  P / H & sentences & entail & contradict & neutral \\ [0.5ex] 
 \hline
 premise & Boy ate an apple. &  -  &  -  & - \\ 
 hypothesis & Kid ate fruits. & \checkmark &   &  \\
 hypothesis & Boy ate an banana. &  & \checkmark & \\
 hypothesis & Tom ate an apple &  &  & \checkmark \\
 hypothesis & Boy ate a Fuji apple. &   &  & \checkmark \\
 %\hline
 %premise & Facebook's IPO electrified the general public. &  -  &  -  & - \\ 
 %hypothesis & Facebook went public. & \checkmark   &  &  \\
 %hypothesis & General Electric went public. &  &   & \checkmark \\
 %hypothesis & People ignored Facebook's IPO. &  & \checkmark & \\
 \hline
\end{tabular}
\end{table}

@@

\taoli{The example needs to be changed accordingly.}

>>>>

\subsection{Natural Language Inference}
\label{sec:languageInference}
Natural Language Inference (NLI)~\cite{DaganRothSammons2013} is an important machine understanding task in NLP.
The goal of NLI is to predict the relationship between a premise (\textbf{P}) sentence and a hypothesis (\textbf{H}) sentence.
The prediction falls in one of three categories: \emph{entailment} (\textbf{E}), \emph{contradiction} (\textbf{C}), \emph{neutral} (\textbf{N}).
A simple example is shown in Table~\ref{tab:NLI}.
In this case, the premise is ``Boy ate an apple''.
The hypothesis statement ``Kid ate fruits.'' can be concluded from the premise. Therefore, the relationship between premise and hypothesis is \emph{entailment}. However, we should note that such a relationship is not necessarily reversible. Since the concept ``fruit'' is less restrictive than that of ``apple'', we can not conclude ``Boy ate an apple'' from the statement ``Kid ate fruit''. The same logic applies to the hypothesis of ``Boy ate a Fuji apple.'', in which the premise neither imply nor oppose the hypothesis, therefore, their relationship is \emph{neutral}.
%\emph{Contradiction} means the textual of hypothesis opposes that of premise;
%and \emph{Neutral} implies no conclusion can be made on either \emph{Entailment} or \emph{Contradiction}.

\begin{table}[htbp]
\label{tab:NLI}
\centering
\caption{An illustration of natural language inference.}
 \begin{tabular}{c | c c c c}
 \hline
  P / H & sentences & entail & contradict & neutral \\ [0.5ex]
 \hline
 premise & Boy ate an apple. &  -  &  -  & - \\
 hypothesis & Kid ate fruits. & \checkmark &   &  \\
 hypothesis & Boy ate an banana. &  & \checkmark & \\
 hypothesis & Tom ate an apple. &  &  & \checkmark \\
 hypothesis & Boy ate a Fuji apple. &   &  & \checkmark \\
 %\hline
 %premise & Facebook's IPO electrified the general public. &  -  &  -  & - \\
 %hypothesis & Facebook went public. & \checkmark   &  &  \\
 %hypothesis & General Electric went public. &  &   & \checkmark \\
 %hypothesis & People ignored Facebook's IPO. &  & \checkmark & \\
 \hline
\end{tabular}
\end{table}

@@

### 2018-03-27 paper/task.tex
add attention view
update task

%Through many rounds of discussions with NLP experts,
In such a highly exploratory process, the expert likely will have many ``what if'' type of questions (e.g., what if I make change to this word in the input, what if I force the words to align correctly in the attention), get responsive answers to these queries is essential to maintain the flow of reasoning.
%
Here, we solidify the analysis goals around the support for these type workflow and thought process of the experts.
%
\begin{itemize}
\item Firstly, understanding how perturbation (e.g., replacing words) of the input affect the prediction is the curial part of error analysis. It not only measures the prediction sensitivity, but also help reveal the potential source of the errors in the input (\textbf{T1}).

%Understand how attention affect the prediction.
\item Secondly, study the role and mechanism of attention by examine how variation in the input affect the attention, and how changes in attention affect prediction (\textbf{T2}).
%Study the relationship between model stages (encoder, attention, classifier) and the prediction
So far, we focus on how changes in earlier stage of the model affect later stages or final prediction.

\item However, what if the current prediction is wrong, how can we apply minimal update to the model to produce correct prediction and how would such a modification affect attention.
Therefore, thirdly, we want to explore how model should be updated to produce the desirable change of prediction, and how the attention respond to such a change (\textbf{T3}).
\end{itemize}
%
% Based on the NLP analysis goals, we identify the visualization tasks to achieve these goals.

>>>>

%Through many rounds of discussions with NLP experts,
In such a highly exploratory process, the expert likely will have many ``what if'' type of questions (e.g., what if I make change to this word in the input, what if I force the words to align correctly in the attention), get responsive answers to these queries is essential to maintain the flow of reasoning.
%
Here, we solidify the visualization tasks to facilitate the exploratory workflow and thought process of the domain experts.
%
\begin{itemize}
    \item Firstly, understanding how perturbation (e.g., replacing words with synonymous) of the input affect the prediction is the curial part of error analysis. It not only measures the prediction sensitivity, but also help reveal the potential source of the errors in the input (\textbf{T1}).

    %Understand how attention affect the prediction.
    \item Secondly, study the role and mechanism of attention by examine how perturbation in the input sentence affect the attention, and how changes in attention affect prediction (\textbf{T2}).
    %Study the relationship between model stages (encoder, attention, classifier) and the prediction

    \item So far, we focus on how changes in earlier stage of the model affect later stages or final prediction. However, what if the current prediction is wrong, how can we apply minimal update to the model to produce correct prediction and how would such a modification affect attention.
    Therefore, thirdly, we want to explore how model should be updated to produce the desirable change of prediction, and how the attention respond to such a change (\textbf{T3}).
\end{itemize}
%
% Based on the NLP analysis goals, we identify the visualization tasks to achieve these goals.
